---
title: "stringsAsFactors = HELLNO"
author: "Xinmiao Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stringsAsFactors = HELLNO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply concatenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```

Sometimes, character is read as a factor in a wrong way. The unique values of a factor tends to be different from its length; whereas the unique values of character is equal to its length To detect the factor which are supposed to be character, I develop a function `detect_fct()`. 

```{r}
c <- as.factor(c("yes", "yes", "no", "yes"))
detect_fct(c) #return TRUE
```

```{r}
d <- as.factor(c("Apple", "Banana", "Orange", "Cherry"))
detect_fct(d) #return FALSE
```

To reorder a factor in a decreasing order, I develop function `reorder()`. 

```{r}
e <- as.factor(c(21,23,45,64,12,22,3, 4, 4, 4, 4))
levels(e)
```

```{r}
f <- reorder(e)
levels(f)
```


Usually, we merely want to encode a vector of data as a factor with the order in which they appear in the data. Hence, based on `factor`, I develop a function `factor_asis()`. By `factor_asis()`, we can set levels to the original order.

```{r}
g <- c(1, 2, 3, 14, 0, 7)
factor_asis(g)
```


